### 2010-02-24 sarahDraft.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage[hmargin=3cm, vmargin=3cm]{geometry}
\begin{document}

@@

		All of the models that we looked at were single-species, single-season
models because this is what fit our data, but any of these models can be modified
to take into account multiple seasons or multiple species.

	The most simple way to measure occupancy is to simply divide the number 
of sites where an animal was found to be present by the total number of sites.  
This is called naive occupancy [MacKenzie2006].  It is generally accepted that 
this estimate is too low because animals are not detected at all sites where they
are present, which is the issue of false negatives.  The other occupancy models
attempt to do better than this simple estimate.

	One way to create more accurate models is to use a Hierarchical model to 
think about the parameters that affect occupancy.

	A more complicated model that can be used as a Hierarchical model is Maximum
Likelihood Estimation (MLE).  This is done by finding the probability 
distribution of the observed data as a function of the parameters.  This 
distribution is then said to be the function of the parameters conditional on the
data, and becomes the likelihood function.  Then, the values of the parameters 
that maximize the likelihood function are found, which means that we find the 
parameters that make the data most likely given the underlying model.  
[MacKenzie2006]

	Another way to do this is using a Bayesian model, which we will discuss in 
more detail below.

@@

	\subsection{Bayesian Statistics}
	   	Bayes' Theorem \\
    	Useful for hierarchical modeling: look at section 3.3 in book. \\
    	Can use Markov Chain Monte Carlo (MCMC) to calculate posterior 
			distribution- bimodal issues.

>>>>

\documentclass{article}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage[hmargin=3cm, vmargin=3cm]{geometry}
\usepackage[all]{xy}
\begin{document}

@@

		All of the models that we looked at were single-species, single-season
models because this is what fit our data, but these models can be modified to
be used for multiple seasons or multiple species.

	The most simple way to measure occupancy is to simply divide the number 
of sites where an animal was found to be present by the total number of sites.  
This is called naive occupancy [MacKenzie2006].  It is generally accepted that 
this estimate is too low because animals are not detected at all sites where they
are present, which is the issue of false negatives.  Other occupancy models
attempt to do better than this simple estimate.

	One way to create more accurate models is to use a Hierarchical model as a 
way to think about the parameters that affect occupancy.  This allows us to take
into account that when an animal is not detected, it could either not be there, 
or it could be there and just was not detected.  This could either be due to
a false negative, where the observer did not see the indicator, or it could be 
due to the fact that the animal does occupy the site but there was no indication
of the animal at the site- for example, no track, no vocalization, etc.  \\
Here is a visualization of how the Hierarchical model works:
	\begin{displaymath}
		\xymatrix{          	   &                                  		 &                               		&\text{Site} \ar[dl] \ar[dr] & \\
				            	   &                                  		 & \text{Occupied} \ar[dl] \ar[d]		&                            & \text{Unoccupied} \ar[d] \\
				            	   & \text{Presence Indicated} \ar[dl] \ar[d]& \text{Presence Not Indicated} \ar[d]&                            & \text{Not Detected} \\
				   \text{Detected} & \text{Not Detected}              		 & \text{Not Detected}                  &                            &   }
	\end{displaymath}

From this diagram, you can see that we need separate parameters for the 
probability that a site is occupied, the probability that presence is indicated
given that a site is occupied, and the probability of detection given that 
presence is indicated and the site is occupied.  We call these \(\psi\),
\(\theta\), and \(p\), respectively.

@@

	\subsection{Bayesian Statistics}

	   	Bayes' Theorem: \(P(\theta|X) = P(X|\theta)P(\theta) / P(X)\) \\

	Usually, the mean of the posterior distribution is then used as a point 
estimate for the parameter.

Therefore, the posterior distribution \( \propto \) the likelihood times the prior.
    	Useful for hierarchical modeling: look at section 3.3 in book. \\
    	Can use Markov Chain Monte Carlo (MCMC) to calculate posterior 
			distribution- bimodal issues.
